---
title: "solutions session 7"
format: html
---

## Exercises 1

1. Get the table with 2023 opinion polling for the next United Kingdom general election from <https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_2024_United_Kingdom_general_election&oldid=1232596394>

```{r}
library(tidyverse)
library(rvest)
# 1. Request & collect raw html
html <- read_html("https://en.wikipedia.org/w/index.php?title=Opinion_polling_for_the_2024_United_Kingdom_general_election&oldid=1232596394")

# 2. Parse
opinion_table <- html |>
  html_elements(".wikitable") |> 
  html_table() |>                
  pluck(2)                    
```


2. Wrangle and plot the data opinion polls

```{r}
# 3. Wrangle
opinion_tidy <- opinion_table |> 
  pivot_longer(Con:Others, names_to = "party", values_to = "result") |> 
  mutate(result_pct = as.numeric(str_extract(result, "\\d+(?=%)")),
         date_clean = str_extract(Datesconducted, "\\d{1,2} [A-z]{3}"),
         date = lubridate::dmy(paste(date_clean, 2024))) |> 
  filter(!is.na(result_pct),
         # some dates are parsed wrong and are in the future
         date < "2024-07-03")

# Plot!
opinion_tidy |> 
  ggplot(aes(x = date, y = result_pct, colour = party)) +
  geom_line()
```

## Exercises 2

1. For extracting text, `rvest` has two functions: `html_text` and `html_text2`. Explain the difference. You can test your explanation with the example html below.

```{r}
html <- "<p>This is some text
         some more text</p><p>A new paragraph!</p>
         <p>Quick Question, is web scraping:

         a) fun
         b) tedious
         c) I'm not sure yet!</p>" |> 
  read_html()
```

- You should almost always use `html_text2`, as the help page says (`?html_text2`)
- `html_text2`, put briefly, extracts text as the browser would display it
- `html_text`, retains the original line breaks and extra space in the raw html content
- You can easily test that with

```{r}
html_text(html)
html_text2(html)
```

The superfluous line breaks and extra space from the html code is not displayed

2. How could you convert the `links` objects so that it contains actual URLs?

The links are all relative to https://en.wikipedia.org. So we need to add that part to the links:

```{r}
# using base R's paste
paste0("https://en.wikipedia.org", links)
# Or str_c from stringr
str_c("https://en.wikipedia.org", links)
```

3. How could you add the links we extracted above to the `pm_table` to keep everything together?

- We can use `left_join` for that. First, we need to put `links` int a table. But we already did to construct `links_df`
- The primary key in pm_table would be `Prime ministerOffice(Lifespan)` and the foreign key is just `name`
- This is basically all I wanted to know here, but if you want to go through with it, we need to clean up  `Prime ministerOffice(Lifespan)` a bit first

```{r}
# 1. Request & collect raw html
html <- read_html("https://en.wikipedia.org/w/index.php?title=List_of_prime_ministers_of_the_United_Kingdom&oldid=1166167337") # I'm using an older version of the site since some just changed it

# 2. Parse
pm_table <- html |> 
  html_element(".wikitable:contains('List of prime ministers')") |>
  html_table() |> 
  as_tibble(.name_repair = "unique") |> 
  filter(!duplicated(`Prime ministerOffice(Lifespan)`))

# 3. No wrangling necessary
pm_table

pm_links <- tibble(name = title, link = links)
pm_table_clean <- pm_table |> 
  mutate(pm = str_extract(`Prime ministerOffice(Lifespan)`, ".+\\["),
         pm = str_remove(pm, "\\["))

pm_table_clean |> 
  left_join(pm_links, by = c("pm"= "name"))
```

## Exercises 3

We might be interested in whether a purchase was verified or not

1. Extract that information from the first review page

We have already extracted the reviews from the first page, so we can work with them directly:

```{r}
verified <- reviews |> 
  html_elements("[data-hook=\"avp-badge\"]") |> 
  html_text2()
```

2. Add the variable verified to the `parse_review` function

```{r}
parse_review <- function(r) {
  rating <- r |> 
    html_element(".review-rating") |> 
    html_text2()
  
  title <- r |> 
    html_elements(".a-letter-space+span") |> 
    html_text2()
  
  date <- r |> 
    html_elements(".review-date") |> 
    html_text2()
  
  text <- r |> 
    html_elements(".review-text-content") |> 
    html_text2()
  
  helpful <- r |> 
    html_elements("[data-hook=\"helpful-vote-statement\"]") |> 
    html_text2()
  
  if (length(helpful) == 0) {
    helpful <- NA_character_
  }
  
  verified <- reviews |> 
    html_elements("[data-hook=\"avp-badge\"]") |> 
    html_text2()
  
  tibble(date, rating, helpful, verified, title, text)
}
```


3. Create `all_reviews` again, but with the verified variable this time

We do not actually have to change anything about the last part. As the `parse_review` is called inside `get_reviews_from_page`, it automatically uses the right code for us

```{r}
get_reviews_from_page <- function(link) {
  html <- read_html(link)
  
  reviews <- html |> 
    html_elements("[id^=customer_review-]")
  
  map(reviews, parse_review) |> 
    bind_rows()
}

all_reviews <- map(links, get_reviews_from_page) |> 
  bind_rows()
glimpse(all_reviews)
```

## Exercises 4

1. Get the author, publication datetime, headline and text from this site: <https://www.cnet.com/tech/services-and-software/facebook-hopes-to-normalize-idea-of-data-scraping-leaks-says-leaked-internal-memo/> (hint: it works in a very similar way, but you have to apply one extra data wrangling step)

```{r}
# 1. Request & collect raw html
html <- read_html("https://www.cnet.com/tech/services-and-software/facebook-hopes-to-normalize-idea-of-data-scraping-leaks-says-leaked-internal-memo/")

# 2. Parse
json_string <- html |> 
  rvest::html_element("[type=\"application/ld+json\"]") |>
  rvest::html_text() 

# 3. wrangling (part 1)
data <- jsonlite::fromJSON(json_string, simplifyVector = FALSE)

data_article <- pluck(data, "@graph", 1)

headline <- data_article$headline
author <- data_article[["author"]][[1]][["name"]]
date <- data_article$datePublished |> 
  lubridate::as_datetime()

text <- html %>%
  rvest::html_elements(".c-pageArticle_content p") %>%
  rvest::html_text2() %>%
  paste(collapse = "\n")

article_df <- tibble(date, author, headline, text)
```



