---
title: "solutions"
format: html
---

## Exercises 1

1. Get the table with 2023 opinion polling for the next United Kingdom general election from <https://en.wikipedia.org/wiki/Opinion_polling_for_the_next_United_Kingdom_general_election>
2. Wrangle and plot the data opinion polls

## Exercises 2

1. For extracting text, `rvest` has two functions: `html_text` and `html_text2`. Explain the difference. You can test your explanation with the example html below.

```{r}
html <- "<p>This is some text
         some more text</p><p>A new paragraph!</p>
         <p>Quick Question, is web scraping:

         a) fun
         b) tedious
         c) I'm not sure yet!</p>" |> 
  read_html()
```

2. How could you convert the `links` objects so that it contains actual URLs?
3. How could you add the links we extracted above to the `pm_table` to keep everything together?

## Let's scrape! Exercise

I started the code below, now it's your turn to finish it:

```{r}
#| eval: false
# 1. Request & collect raw html
html <- read_html("https://www.ic2s2.org/program.html")

sessions <- html |> 
  html_elements(".nav_list")

# 2. Parse
talks <- sessions |> 
  html_elements("li")

talks_titles <- talks |> 
  html_elements("") |> 
  html_text()

talks_speaker <- talks |> 
  html_elements("") |> 
  html_text()

talks_authors <- talks |> 
  html_elements("") |> 
  html_text()
``` 

## Exercises 3

1. Can you find older iterations of the ECPR conference? How would you scrape these programmes?

