---
title: "solutions day 1"
format: html
---

## Exercises 2

1. Fork the course repository on GitHub
2. Clone the course repository to your computer
3. In the folder "participants" copy the file "participants.csv" and name the copy "participatns_YOUR_NAME" 
4. Fill in your details (if you feel uncomfortable sharing your details with the class and on the public GitHub site, just add "-" to some or all columns)
4. Commit and push your changes
5. Make a pull request to the main repo
6. Optional: After I have merged your pull request, you can set your fork to private and add me and Marius as collaborators (you will submit your homework that way)

## Exercises 3

1. Run ggplot(data = mpg). What do you see and why?

- an empty plot
- ggplot creates the "canvas" of the plot with the right dimensions. But since we added no aestetics or layers, the canvas remains empty

2. In the function `pb_collect()` from `paperboy`, what do the arguments `ignore_fails` and `connections` do?

- to check this, use the command `?paperboy::pb_collect()`
- `ignore_fails` ignores individual connections failing and moves on to the next URL to download data from
- `connections` sets the number of connections opened at once

3. Write a function that takes a numeric vector of miles per gallon consumption data and transforms it to kilometer per liter. If anything other than a numeric vector is entered, the function should display an error (hint: see ?stop).

```{r}
# reusing the formula from the section 'The pipe"
mpg_converter <- function(mpg) {
  if (!is.numeric(mpg)) {
    stop("only numeric vectors allowed")
  }
  kpl = mpg * 0.4251
  return(kpl)
}
mpg_converter(c(20, 25, 30))
mpg_converter(c("twenty-five"))
```


4. In the code below, check the sizes of the intermediate objects with `object.size()`.

```{r}
file_link <- "https://raw.githubusercontent.com/shawn-y-sun/Customer_Analytics_Retail/main/purchase%20data.csv"
df <- read.csv(file_link)
filtered_df <- df[df$Age >= 50,]
aggregated_df <- aggregate(filtered_df$Quantity, by = list(filtered_df$Day), FUN = sum)
names(aggregated_df) <- c("day", "total_quantity")
aggregated_df[order(aggregated_df$total_quantity, decreasing = TRUE)[1:5],]
```

```{r}
object.size(df)
object.size(filtered_df)
object.size(aggregated_df)
```


5. How could the code above be improved if you only want the final result, the code should be readable and you care about memory usage?

It could be improved by not saving intermediate objects, but using the pipe instead.
I also convert it to tidyverse code, but that is optional:

```{r}
library(tidyverse)
final_df <- file_link |> 
  read.csv() |> 
  filter(Age >= 50) |> 
  group_by(Day) |> 
  summarise(total_quantity = sum(Quantity)) |> 
  arrange(-total_quantity)
object.size(final_df)
head(final_df)
```

Another alterantive would be to wrap this in a function, as intermediat object inside the function environment are destroyed and only the final object (which should be wrapped in `return()`) is added to the main environment/memory.

```{r}
top_quantity_days <- function(file_link) {
  df <- read.csv(file_link)
  filtered_df <- df[df$Age >= 50,]
  aggregated_df <- aggregate(filtered_df$Quantity, by = list(filtered_df$Day), FUN = sum)
  names(aggregated_df) <- c("day", "total_quantity")
  aggregated_df[order(aggregated_df$total_quantity, decreasing = TRUE)[1:5],]
}
top_quantity_days("https://raw.githubusercontent.com/shawn-y-sun/Customer_Analytics_Retail/main/purchase%20data.csv")
```


## Exercises 5

1. Write a Python function that takes a numeric vector of miles per gallon consumption data and transforms it to kilometer per liter. Use the function from within R

```{r}
library(reticulate)
```


```{python}
def mpg_converter(mpg):
    kpl = mpg * 0.4251
    return kpl
```


```{r}
library(reticulate)
py$mpg_converter(20)
```

Note: the function only takes a single numeric value. To make it take a longer vector is a little more advanced:

```{python}
def mpg_converter(mpg_list):
    kpl_list = []
    for mpg in mpg_list:
        kpl = mpg * 0.4251
        kpl_list.append(kpl)
    return kpl_list
```

```{r}
py$mpg_converter(c(20, 25, 30))
```

2. Use the function in Python, but using a vector defined in R

```{r}
mpg <- c(20, 25, 30)
```


```{python}
mpg_converter(r.mpg)
```


