---
title: "solutions session 7"
format: html
---

---
title: "solutions day 1"
format: html
---

## Exercises 1

1. `httr2` has several more functions to customize how a request is performed. What do these functions do

- `req_throttle`: makes sure to not exceed rate limits of the API by stopping and waiting at a given rate
- `req_error`: controls how errors should be handled. Can change the interpretation of http codes and display a different error body using a function
- `req_retry`: if a request fails, this function controls whether, after how long and how often the request is retired

2. You might want to add more information to the data.frame. Adapt the function parse_response to also extract: apiUrl, lastModifiedDate, pillarId

```{r}
parse_response <- function(res) {
  tibble(
    id = res$id,
    type = res$type,
    time = lubridate::ymd_hms(res$webPublicationDate),
    headline = res$webTitle,
    text = read_html(pluck(res, "blocks", "body", 1, "bodyHtml")) |> html_text2(),
    api_url = res$apiUrl,
    pillar_id = res$pillarId,
    last_modified = pluck(res, "blocks", "body", 1, "lastModifiedDate")
  )
}
parse_response(res)
```


3. Make your own request to the API with a different search term

```{r}
library(httr2)
library(tidyverse, warn.conflicts = FALSE)
req <- request("https://content.guardianapis.com") |>  
  req_url_path("search") |>                            
  req_method("GET") |>                                 
  req_timeout(seconds = 60) |>                            
  req_headers("User-Agent" = "httr2 guardian test") |> 
  req_url_query(                                       
    q = "my own request",  # <- this is where the search term goes
    "show-blocks" = "all"
  ) |> 
  req_url_query(
    "api-key" = Sys.getenv("GUARDIAN_KEY")
  )
  
# we can also directly parse this
req |> 
  req_perform() |> 
  resp_body_json() |> 
  pluck("response", "results") |>
  map(parse_response) |> 
  bind_rows()
```


4. Request page 2 from the API

The documenation at <https://open-platform.theguardian.com/documentation/> tells us the following:

> In order to page through the results, you can add the page keyword to your query.

```{r}
req <- request("https://content.guardianapis.com") |>  
  req_url_path("search") |>                            
  req_method("GET") |>                                 
  req_timeout(seconds = 60) |>                            
  req_headers("User-Agent" = "httr2 guardian test") |> 
  req_url_query(                                       
    q = "my own request",
    "show-blocks" = "all",
    page = 2  # <- this is where we can add the page parameter
  ) |> 
  req_url_query(
    "api-key" = Sys.getenv("GUARDIAN_KEY")
  )
```


## Exercises 2

To get more information about an MP, we can use the endpoint "/api/Members/{id}/Biography"

1. Search for an MP you are interested in with the function above and use the id on the documentation website with "Try it out"

- you can just search the page <https://members-api.parliament.uk/index.html> for "/api/Members/{id}/Biography"
- the ID of John Major is 119, which you can use as an example
- It produces this call

```
curl -X 'GET' \
  'https://members-api.parliament.uk/api/Members/119/Biography' \
  -H 'accept: text/plain'
```

2. Copy the Curl call and translate it into `httr2` code

```{r}
library(httr2)
curl_translate("curl -X 'GET' \
  'https://members-api.parliament.uk/api/Members/119/Biography' \
  -H 'accept: text/plain'")
```

3. Wrangle the `governmentPosts` returned in the data into a tabular format

```{r}
resp <- request("https://members-api.parliament.uk/api/Members/119/Biography") |> 
  req_method("GET") |> 
  req_headers(
    accept = "text/plain",
  ) |> 
  req_perform()

# you could check which content type is returned
resp_content_type(resp)


major <- resp |> 
  resp_body_json()

government_posts <- purrr::pluck(major, "value", "governmentPosts")
dplyr::bind_rows(government_posts)
```

Bonus:

4. Write a function which lets you request information given an ID and which wrangles the results

```{r}
get_government_posts <- function(id) {
  
  resp <- request("https://members-api.parliament.uk") |> 
    req_url_path("/api/Members", id, "Biography") |> 
    req_method("GET") |> 
    req_headers(
      accept = "text/plain",
    ) |> 
    req_perform()
  
  # you could check which content type is returned
  if (resp_content_type(resp) != "application/json") {
    stop("incorrect data type returned from API")
  }
  
  major <- resp |> 
    resp_body_json()
  
  government_posts <- purrr::pluck(major, "value", "governmentPosts")
  dplyr::bind_rows(government_posts)
}
get_government_posts(119)
```


5. Two more interesting endpoints are "/api/Posts/GovernmentPosts" and "/api/Posts/OppositionPosts". What do they do and how can you request data from them


- "/api/Posts/GovernmentPosts": Returns a list of government posts.

```{r}
curl_translate("curl -X 'GET' \
  'https://members-api.parliament.uk/api/Posts/OppositionPosts?departmentId=51' \
  -H 'accept: text/plain'")

request("https://members-api.parliament.uk/api/Posts/OppositionPosts") |> 
  req_method("GET") |> 
  req_url_query(
    departmentId = "51",
  ) |> 
  req_headers(
    accept = "text/plain",
  ) |> 
  req_perform()
```



- "/api/Posts/OppositionPosts": Returns a list of opposition posts.


```{r}
curl_translate("curl -X 'GET' \
  'https://members-api.parliament.uk/api/Posts/GovernmentPosts?departmentId=51' \
  -H 'accept: text/plain'")

request("https://members-api.parliament.uk/api/Posts/GovernmentPosts") |> 
  req_method("GET") |> 
  req_url_query(
    departmentId = "51",
  ) |> 
  req_headers(
    accept = "text/plain",
  ) |> 
  req_perform()
```

- Now what is not immediately clear about these two endpoints is how to obtain the department ID. However, there is another endpoint, "/api/Posts/Departments/{type}" which returns a list of departments, including their ID. I made a request and looked through the list to find the ID for the home office (51). This is what I used above. But you can of course use any other ID here.


